---
title: "The M&M Color Study"
author: "Brent L Daniel"
date: "February - March, 2018"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

````{r, echo=FALSE}

````

## M&Ms are colorful, fun and make for great data

I bet when you open a bag of M&Ms, you look at a few of them in your hand, admire the pretty color variation that exists, and just **eat them**. Yes, I used to do that, too. But in March, 2014, I spilled bag of M&Ms onto my desk and noticed something odd. I used a tool from my days as a Six Sigma Black Belt and created a Pareto Chart, and snapped this picture:

((replace with picture with only one Brown M&M))

Wait a minute! Only one brown? Why aren't they more even?

To make a long story short, I posted this on Facebook, and started routinely making "M&M Pareto Charts" and posting them to an album. A friend joined in and posted ones with Peanut M&Ms. As of today, we have logged ((make dynamic)) bags of M&Ms in this manner.

And recently, I compiled the posts into a data set. 

## M&Ms and the questions I might ask about colors

This story is about data, how to manipulate it and analyze it with data visualization tools, and in so doing explore many aspects of the colors of M&Ms! 

I became fascinated by what I was seeing because I saw so much variation. I mean, the naive person might just assume that every bag had (roughly) the same amount of each color. You could imagine a manufacturing process that did that, right? But could you imagine a manufacturing process that wasn't predictable in terms of the distribution of colors? That's what I was seeing!

As I looked at these pictures, I came to be curious about a number of things:

* how often is there only one of one color?
* how does a bag have only five of the six colors?
* does blue really dominate the counts, as it seems to?
* how does the color distribution differ by size of bag?
    + are smaller bags more variable? 
    + are smaller bags more likely to have missing colors?
    + what's different about the largest bags?
* do Peanut and Plain M&Ms have similar color distrubtions?

## maybe add pictures for each question?

### Why Pareto Charts of M&Ms

- pretty
- category is its own label (show bag and distribution vs. picture)



## Preparing the data for analysis
# Tidy set, untidy set, pros and cons

- kind of data: categorical, nominal

## 
```{r}
vals <- c(2, 4, 1,2,4,3)
colorsm <- c("Yellow","Blue","Green","Brown","Orange","Red")
dft<-data.frame(vals,colorsm)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
